### 2016-05-18 WRITEUP/table_medianLens.tex
done lens modeling sec

\begin{deluxetable*}{lcc}[!htbp]
\tabletypesize{\scriptsize}
\tablecolumns{3}
\tablecaption{Lens parameters from preliminary models}
\tablehead{
\multicolumn{2}{c}{Parameters} &
\colhead{Median values} % weighted-mean of median of each
}
\startdata
Offset in RA    & (\arcsec)   &  0.004$\pm$0.027\\
Offset in Dec    & (\arcsec)   & 0.003$\pm$0.027\\
Axial Ratio      &             & 0.56$\pm$0.16\\
Position Angle   & (deg)       & 103$\pm$22\\
Einstein Radius  & (\arcsec)   & 1.833$\pm$0.002\\
\enddata
\label{tab:lens}
\tablecomments{ Parameters describing the foreground lens are
obtained based on the preliminary models (see text for details).
All angular offsets are with respect to
$\alpha$\,=\,11$^{\rm h}$31$^{\rm m}$51\fs44,
$\delta$\,=\,-12\degr31\arcmin58\farcs3 (J2000).
The corresponding masses
within the Einstein radii is $M(\theta$\,\,$<$\,\,$\theta_\textrm{E})$\,=\,(7.47\,$\pm$\,0.02)\,$\times$\,10$^{11}$\,\,\Msun.}
\end{deluxetable*}