### 2016-05-25 WRITEUP/table_magFactor.tex
updated SED section with SFR and comparison to C13 sameple, but LIR and FIR/IR fraction, etc

\begin{deluxetable}{lcc}[!htbp]
\tabletypesize{\scriptsize}
\tablecolumns{3}
\tablecaption{Magnification factors of various kinematic components in \bco}
\tablehead{
\colhead{Channel} &
\colhead{Source 1 $\mu_{\rm L}$} &
\colhead{Source 2 $\mu_{\rm L}$}
}
\startdata
126$-$130 & 7.2 $\pm$ 5.6 & 6.7 $\pm$ 2.5 \\ [0.6ex] %which is source 1 and 2 then?
131$-$135 & 7.6 $\pm$ 1.6 & \\ [0.6ex] % would this be source 1 or 2?
136$-$140 & 8.7 $\pm$ 2.0 & \\ [0.6ex]
141$-$145 & 4.1 $\pm$ 0.9 & \\ [0.6ex]
146$-$150 & 4.2 $\pm$ 0.6 & \\ [0.6ex]
151$-$155 & 4.3 $\pm$ 2.4 & \\ [0.6ex]
156$-$160 & 3.1 $\pm$ 0.9 & \\ [0.6ex]
weighted average & 4.4 & \\ [0.6ex]
median & 5.5 &
\enddata
\label{tab:model}
\tablecomments{Source 1 is RXJ1131 and source 2 is its companion. See text for details.}
\end{deluxetable}