### 2016-06-19 WRITEUP/table_magFactor.tex
formatting

\begin{deluxetable}{lcc}[!htbp]
\tabletypesize{\scriptsize}
\tablecolumns{3}
\tablecaption{Magnification factors of various kinematic components in \bco}
\tablehead{
\colhead{Velocity Range (\kms)} & % see 22May16/intensity30Apr16.spec; or channel map in paper
\colhead{Source 1 $\mu_{\rm L}$} &
\colhead{Source 2 $\mu_{\rm L}$}
}
\startdata
-366\,$-$\,-258 & 3.1 $\pm$ 0.9 & \\ [0.5ex]
-237\,$-$\,-151 & 4.3 $\pm$ 2.4 & \\ [0.5ex]
-129\,$-$\,-43  & 4.2 $\pm$ 0.6 & \\ [0.5ex]
-21.5\,$-$\,65    & 4.1 $\pm$ 0.9 & \\ [0.5ex]
86\,$-$\,172        & 8.7 $\pm$ 2.0 & \\ [0.5ex]
194\,$-$\,280       & 7.6 $\pm$ 1.6 & \\ [0.5ex]
301\,$-$\,388       & 7.2 $\pm$ 5.6 & 6.7 $\pm$ 2.5 \\ [0.5ex]
weighted average & 4.4 & \\ [0.5ex]
median & 5.5 &
\enddata
\label{tab:model}
\tablecomments{Velocity is taken from the center of each (native) channel
without any binning. Each row corresponds to a channel slice used for
lens modeling. Source 1 is RXJ1131 and source 2 is its companion. See text for details. }
\end{deluxetable}