### 2016-11-05 WRITEUP/response.md
update redshift

We thank the referee for providing constructive and thoughtful
comments to improve the manuscript. We have addressed all the concerns
raised by the referee and details can be found in the revised version
-- please see attached. A brief description on how we incorporated the
changes can be found below.

1a) We have changed the title to ".... of the strongly lensed quasar host galaxy RXS\,J1131-1231"

1b) To minimize potential confusion, we have changed "dynamically resolved" to "kinematically resolved".

1c) We have updated our redshift (and re-plotted the figure) based on the results from fitting a double-Gaussian to the de-lensed intrinsic line profile of our target (Fig. 5).

1d) We have changed "dynamical information" to "kinematic information".
- reconstruct the intrinsic gas dynamics --> reconstruct the intrinsic line profile and source-plane velocity structure

We have rephrased the 2nd paragraph of Sec 4.1.2 to clarify our message to the reader.