### 2016-11-22 WRITEUP/response.md
revised with DR's comments

We thank the referee for providing constructive and thoughtful
comments that helped us to improve this manuscript. We have addressed all the concerns raised by the referee in the revised version of the manuscript. Our response can be found after each of the referee's points below.

COMMENTS IN DETAIL:

1) TOO MANY "DYNAMICS" AND "DYNAMICAL".

You may be mis-using the word "dynamical" in some places.
"Dynamics" refers to the effects of a force, or forces, in 
causing motions, or in reaching equilibrium.

So if you are mainly talking about velocities, it would be 
better to say "velocities" or "kinematics" instead of "dynamics".

1a): CHANGE THE TITLE : DROP THE WORD "DYNAMICS" :

The Title says the paper is about the "dynamics" of the molecular gas in the 
quasar host galaxy.

Most of the paper, however, is not about the "dynamics" at all, 
but rather about the CO detections,
the VLA continuum map (synchrotron emission, not molecular gas), 
the spectral energy distribution in the optical, near-IR, mid-IR, and far-IR continuum,
and the gravitational lens model.

I suggest that you consider changing the title, to better describe the
contents of the paper, maybe something like:

"CO-Line, Radio Continuum, Optical-to-Far-IR Observations, and Lens Models of
the Host Galaxy of the Quasar RXS J1131-1231"

or maybe:

"Interpretation of CO, Radio, Optical, and IR Continuum 
observations of the Strongly Lensed Host Galaxy of the 
Quasar RXS J1131-1231"

or simply:

"CO, Radio, Optical, and IR results on 
the Lensed Host Galaxy of the Quasar RXS J1131-1231".

# RESPONSE:
1a) To take the referee's point into account, we have changed the title to "Molecular gas kinematics of the strongly lensed quasar host galaxy RXSJ1131-1231". We agree that we study more than just the CO gas of the science target in this paper; however, a large fraction of the paper is based on our new CO data and the analysis (including lens modeling).

1b) DROP "DYNAMICALLY RESOLVED":

Sect.3.1, paragraph 1:

"The emission is spatially and dynamically resolved...as shown in Fig. 1 "

"Dynamically resolved" sounds strange. 
The reader might confuse it with the dynamic range.
Maybe change to :

"The line emission is resolved, in both velocity (Fig. 1) and 
in its spatial distribution (Fig. 2).

# RESPONSE:
We have addressed this comment by changing "dynamically resolved" to "kinematically resolved". to minimize potential confusion.

1c) DROP "DYNAMICAL CENTER":

Fig.3, caption:
"...z = 0.6537, which corresponds to the dynamical center of CO(2-1)."

Change to :

"..which corresponds to zero velocity in the CO(2-1) spectrum (Fig.1)."

# RESPONSE:
We have taken into account the referee's comment and have updated the CO redshift based on the results from fitting a double-Gaussian to the de-lensed intrinsic line profile of our target (Fig. 5), and re-plotted all the relevant figures with the updated velocity scale. We have also changed our wording in fig. 1 caption "dynamical center" to "line center".

1d): DROP "RECONSTRUCT THE DYNAMICS" :

Sect.4.1, para.1:

Too many "dynamics" in this paragraph.

"...the high spectral resolution .. provides dynamical information 
on scales smaller than the beam ..."

Maybe change to "velocity information" or "kinematic information".

"...we reconstruct the intrinsic gas dynamics by carrying out 
lens modeling over different channel slices..."

You are not reconstructing the "dynamics", 
you are reconstructing the intrinsic spectral profile, by separately 
deriving the lens magnifications in seven different velocity channels.

As you say in the Abstract, this shows that the asymmetric line profile
(red peak stronger than blue peak) is actually symmetric.

I don't think this is reconstructing the "dynamics."
The "dynamics" stays the same --- rotation in a gravitational force field.

You have just reconstructed the line shape, that's all.

Contrary to what you say in Sect.4.1, this is actually not 
"similar" to the "differential lensing ...across the HST V, I and H bands (Claeskens+ 2006)

They talk about different magnification due to varying sizes of different temperature regions.
In the Claeskens reconstruction, the blue light (redshifted to V band) comes from young hot stars 
in a nuclear starburst ring --- see their sect.5.2, point 4.

The V-band source is smaller (more compact) than the cooler regions in I and H bands,
in the larger spiral disk farther from the nucleus.

So, they interpret the different magnifications as due to different sizes in the source plane.

This is not the same as magnification varying across the line profile,
due to different source-plane positions, as you are claiming.

# RESPONSE:
We agree with the referee's comment and have changed "dynamical information" to "kinematic information", "reconstruct the intrinsic gas dynamics" to "reconstruct the intrinsic line profile and source-plane velocity structure".

To address the referee's comment on the paragraph (2nd paragraph of Sec 4.1.2.) regarding the differential lensing effect in CO and that in the optical emission (Claeskens+06), we have re-written this paragraph to clarify what we seek to convey -- the difference in magnification factor between CO gas across all channels (µ ~ 5.5) versus those found in the optical emission indicates different morphology and spatial distribution between the CO gas, and the stellar population of different ages.

2) FIX THE REDSHIFT. IT SHOULD BE 0.6541 .

Your beautiful CO(2-1) spectrum gives the best measurement 
so far of the qso host galaxy redshift,
more precise than the optical spectra.

For this reason, you should make a little extra effort to get it exactly right.

2a. OBSERVED FREQUENCY, NOT LO FREQUENCY.

Sect.2.1, para. 1:

You say nu_LO = 139.256 GHz.
But this is your observed frequency, not the LO frequency.
It corresponds to CO(2-1) redshifted to 0.6555.

What probably happened was this:

a) You say that you first observed at CARMA (Feb.2014).
At CARMA, your sensitivity was only good enough to see the red peak, and not the blue one.
From the CARMA data, you derived the wrong redshift of z = 0.6555.

b) You say that you then observed at IRAM (later, in Dec.2014 and Feb.2015).
Because you already had the (incorrect) redshift from CARMA, 
you set the IRAM observing frequency to the wrong redshift of 0.6555 (139.256 GHz),

Then, with the better sensitivity at IRAM, you discovered that the true line profile 
was actually centered closer to z = 0.6537 (139.4074 GHz).

c) Later on, you re-reduced the data, this time with zero velocity corresponding 
to z = 0.6537, for all the figures in this paper.

d) But your Fig.1 (CO 2-1 spectrum), and especially, 
Fig.5 (both spectra, yellow and blue histograms),
show clearly that the line centroid is at +75 km/s relative to zero velocity.

e) At redshifted CO(2-1), there are 2.15 km/s per MHz, so the true line centroid 
is 34.88 MHz lower than your assumed zero frequency, and 
so the true line centroid is at 139.373 GHz.

f) This corresponds to z = 0.6541, as the true redshift of the host galaxy.

g) SUGGESTED CHANGES :

Sect.3.1, para.1: 
"...confirming the redshift at z = 0.6537..."

--- change to:
"...showing that the line centroid is at a redshift of z = 0.6541..."

Fig.3 caption:
"The velocity scale is with respect to z=0.6537, which corresponds to the dynamical center..."

--- change to:
"The zero of the velocity scale is at z=0.6537, which is close to the 
true centroid of the CO(2-1) line at z=0.6541 ...

Sect.4.5 para.2:

---Possibly re-do the fit, relative to a zero velocity at z = 0.6541, as this will change
the V_max in your p-v diagram (Fig.9, right panel).

--- Or else just explain that the V_0 in your Eq(1) comes out to +28 km/s (relative to z=0.6537)
because of the very coarse sampling of your seven 108 km/s-wide bins.

The true line center (relative to z=0.6537) would be at V_0 = +75 km/s if you had
used the full-resolution spectrum (Fig. 1 or Fig.5), rather than the heavily binned channels.

# RESPONSE:
We have corrected nu_LO --> nu_obs.

We agree with the referee and have re-fitted a doubled Gaussian to the de-lensed profile (Fig. 5) to updated the redshift to z=0.65406. We have therefore replotted all the relevant figures (and updated their captions correspondingly), updated column 1 in Table 3 and the labeling in Fig 7, refitted the arctangent model (Fig. 9) and updated parameters in Sec 4.3 correspondingly, and refitted SED with the updated redshift.

3) CO(2-1) FLUX AND SOURCE SIZE :

Sect. 3.1, para.1: CO(2-1) LINE FLUX :

You say the 2-Gaussian fit gives :
red peak = 75.3 mJy (do you mean mJy/beam?), FWHM=179 km/s,
and blue peak = 24 mJy (do you mean mJy/beam?), FWHM = 255 km/s.

This gives 1.06 x 75.3 x 179 /1000 = 14.3 Jy km/s for the red peak,
and gives 1.06 x 24.1 x 255 /1000 = 6.5 Jy km/s for the blue peak.

So the sum of the two peaks is 20.8 Jy km/s, not the value of 24.1 Jy km/s in the text.

Or, are the values you quote for two Gaussian components only the values at the top pixel of the map?
That is, not spatially integrated, while the 24.1 Jy km/s is spatially integrated?
You need to make this clear, both in the text, and in the caption to Fig.1.

If Fig.1 shows only the spectrum at the source maximum, and not the 
spatially integrated spectrum, then you need to change the label on the vertical axis 
to Flux density per beam [ mJy /beam ].

In the re-plot of the CO(2-1) spectrum that you give in Fig.5 (yellow histogram),
the redshifted peak is now 100 mJy, not 75, as in Fig.1.

Does Fig.5 now show the spatially integrated spectrum ? 
What is the difference between this spectrum and the one in Fig.1?

If it is again a spectrum taken at the maximum-intensity point,
then the units on the vertical axis would again have to be [ mJy/beam ].

In Fig.5,it would be good to show what you think is the contribution from the companion galaxy.

# RESPONSE:
The referee is correct and we found a bug in the GILDAS package that affected the statistics and fluxes within the map used for previous line-integrated flux extraction. In the resubmitted version, we have updated the intensity reported for CO(2-1) by summing the fluxes over the FWZI (20.3+/-0.6 Jy km/s). The updated intensity is in agreement with the double-Gaussian fitted to the  extracted spectrum (Fig. 1). Since our lensing-corrected M_gas for both RXJ1131 and the companion are derived from the flux in each channel, this does not change our conclusion.

We attempted to over-plot the companion's spectrum to Fig. 5, as suggested by the referee, but found the figure to be too busy as the plot is showing various spectra with different various linestyles and colors. This main message for this plot is to show that we recover an intrinsically symmetric line profile for the main galaxy which is informative about its gas kinematics.On the other hand, the useful information regarding the companion from the de-lensed spectrum is the gas mass, which we report in the text (and in the table, as suggested by the referee).

Sect. 3.1, para.3: SOURCE SIZE :

I don't think you can measure source size to three significant figures,
especially if you measure the beam-smoothed sizes in the image plane,
and then try to estimate the deconvolved sizes by doing root square
subtraction of the beam sizes (if that's what you did).

An advantage of an interferometer is that you can 
do fits to the source size in the u,v plane, without having
to deconvolve the beam.

Because the whole process is uncertain, 
instead of saying " 5.1 +/- 0.72 by 3.72 +/- 0.66 ",
you could just say 5.1 +/- 0.7 by 3.7 +/- 0.7 .

This is especially true because you say in the next sentence, that the 
emission is not a 2-D Gaussian, so the fitted model is anyway 
not correct.

But an even bigger problem is that these values look like the observed (not deconvolved)
major and minor axis sizes (FWHM) on the integrated map in Fig.2 (upper left).

Are you sure that you have deconvolved these sizes? For a 4.4" x 2" beam?

Please check this.

Even better, try doing a fit in the u,v plane, rather than the sky plane.

Here's another problem with the source size:
In Sect.1, para.3, you say the stellar emission in the host galaxy of RXJ1131 is 
lensed into an Einstein ring of radius 1.83" (diameter = 3.66"). 
So the length of the arc is about 4 or 5 arcsec.

In Table 3, you say the median magnification is a factor of 5.5.

This means the de-magnified CO source DIAMETER is about 1 arcsec.

At z = 0.654, 1 arcsec is 7 kpc.

So the de-magnified RADIUS of the CO source is only 3.5 kpc.

It looks like the intrinsic R_CO = 6 kpc (diameter = 12 kpc)
that you quote elsewhere in the paper is too high:
---Abstract, 
---Sect.4.3, para.3, 
---Sect.4.4, para.1, 
---Sect.5.1.1, para.2,
---Sect.5.5, para.2
---Sect.5.1.1, para.2,
---Sect.6, para.2

# RESPONSE:
Source size: 
We addressed the referee's comment on the *deconvolved* CO source size by  fitting a single Gaussian in uv-plane, which corresponds to 4.5" +/- 0.1" x 3.5" +/- 0.2", consistent with the reported CO size based on image-plane fitting within the uncertainties. We additionally report the source sizes by fitting two Gaussians to provide further evidence that the CO(2-1) emission is  spatially resolved. We have updated Section 3.1 to clarify the discussion.

CO physical size:
The rough estimate of de-magnified CO diameter provided in the referee report is a valid first order estimate, but the true size of the source is complicated by the fact that
1) the lensed emission is not uniformly stretched along the arc (and contains complex structures within the arc), and thus the magnification factor is not uniform across the partial Einstein ring.
2) CO is not subject to dust extinction, therefore its emission may be more extended than optical emission. In the case of RXJ1131, the CO gas is likely to be more extended than the optical emission given that
    i) the deconvolved source size of CO (see above and section 3.1) is larger than the diameter of the Einstein Ring in the optical (~3.6")
    ii) the spatial extent of the CO emission as shown in Fig. 7, where the centroid of the blueshifted emission is further out in the SE compared to the lensed arc in the optical
    iii) the source-plane spatial distribution from the best-fit lens models (Sec 4.3, Fig. 9), where the most extended velocity component is ~6.2 kpc from the line center position, indicating that the CO source size is unlikely to be R ≤ 3.5 kpc.

The R_CO size of ~6 kpc reported throughout the paper is obtained based on the centroid position of the most extended kinematic component obtained from the best-fit lens models. See also Sec 4.3 and our response to point 14 in this referee report for a more detailed explanation regarding the CO size.

4) ERRORS ON CO(3-2):

Sect. 3.2, para 1: 
"We detect CO(3-2) ... in the CARMA data at > 5 sigma significance."(Fig.3).

Something is wrong here. In Fig.3, the peak-to-peak noise 
in 36 km/s channels, is 100 mJy/beam. 
This means 1 sigma is 25 mJy/b.

This disagrees with Sect.2.2, para.3, which says 1 sigma is 
only 13.3 mJy/b in 36 km/s (25 MHz).

So, while the positive-velocity peak in Fig.3 is 4 or 5-sigma,
the negative velocity peak is only 2 or 3-sigma, so not really significant.

Without knowledge of the CO(2-1) spectrum,
the negative-velocity peak at CO(3-2) could not be called a real detection.

Maybe change the first sentence of Sect.3.2 to read:

"We detect CO(3-2) ... in the CARMA data at 5-sigma significance
on the positive-velocity peak. 
Within the noise limits, the CO(3-2) spectrum (Fig.3) may be
consistent with the double-peak profile seen in CO(2-1) (Fig.1)."

Because of the large errors in the CO(3-2) flux, you should not
quote the flux and the errors to three significant figures (Sect. 3.2).

Instead of saying "35.7 +/- 21.9, 
you could just say 36 +/- 22 .

Similarly, in the same paragraph, for the (3-2) / (2-1) ratio,
you could just say 0.7 +/- 0.4 .

In Sect.3.2, para.1, you say the observing frequency is 209 GHz.
So the wavelength is 1.4mm, not 1.5mm. 
(There are two places to fix this).

#RESPONSE
We have considered the referee's point of view, and agree that the most extreme peak to peak noise across two channels is from -50 to +50 mJy/beam. However, this is not representative of the noise across all channels that are being used to derive the line flux. We thus derive the uncertainty based on the RMS obtained from the intensity map.

To address the referee's comment and to demonstrate the significance of our detection that is reported in the paper, we have extracted the peak fluxes and their corresponding uncertainties for the blue and red wing independently. For the blue wing, we find a peak flux density of 5.13 +/- 1.43 Jy km/s /beam in the intensity map, indicating a ≥3sigma detection in this component. For the red wing, we find a peak flux density of 11.45+/-1.99 Jy km/s /beam in the intensity map, indicating a ~6 sigma detection. The SNRs are thus consistent with the reported detection significance of ≥5sigma in the paper which was obtained based on the peak flux density in the intensity map obtained by summing over the full line profile (i.e., including emission in both wings).

Since both CO lines are observed towards the same source, using prior knowledge about the blue wing from the CO(2-1) data to inform about the CO(3-2) emission also appears reasonable.

While the additional figure in I_CO is not significant, we keep this 
to enable reproducible results and to minimize cumulative errors in deriving second order quantities. We only report significant figures when reporting the resulting physical parameters.

We agree with the referee and have updated 1.5mm to 1.4mm.

5) ERRORS ON THE 2MM CONTINUUM FLUX OF RXJ 1131:
Sect. 3.3, para.1:
The wavelengths should be 1.4mm for the CARMA data (not 1.5mm),
and 2.2mm for the IRAM data.

Sect.3.3, para.2: 
"We adopt 0.39 +/- 0.08 mJy as the 2mm continuum ...."

But Table 1 lists "0.400 +/- 0.082 " as the flux.

Both of these entries are wrong.

In Sect.3.3, para.2, you say :

integrated flux = 1.2 +/- 0.2 mJy, 
peak flux = 0.8 +/- 0.09 mJy/b at the lensing galaxy. (round off the 0.799 +/- 0.088).

You subtract a model point source at the lensing galaxy.

This gives a residual of (1.2 - 0.8) = 0.4 mJy.

But since this is a subtraction, you have to add the errors 
in the integrated flux (0.2 mJy)
to the errors in the possible point source (0.09 mJy) to get
the total uncertainty range in the residual.

So the flux of the residual is 0.4 +/- 0.3 mJy, 
not the +/- 0.08 mJy, 
which is only the raw rms in the unsubtracted data.

This means the 2mm continuum "detection" of the lensed arc (Fig.4 upper)
may not be real. 
On the map in Fig.4, it is only a bend in the lowest contour (3-sigma), 
so not really significant.

It might be better to rewrite Sect.3.3, para.2 to say something like:

"We detect 2mm continuum with an integrated flux density of 1.2+/-0.2 mJy,
with a peak flux of 0.80+/-0.09 mJy/beam, centered on the lensing galaxy.
In the lowest contour (3-sigma), there is a hint of possible emission
from the lensed arc."

Instead of calling this the "deblending procedure",
which people might confuse with deconvolution, 
I would just call it "subtraction of the foreground galaxy", 
or "flux removal by subtraction", because it is just simple 
subtraction, not anything else.

# RESPONSE
In the residual map, the peak emission is 0.39 +/- 0.08 mJy. This flux (0.39 mJy) is what we adopted for RXJ1131 at 2mm in the paper. To address the referee's comment, we have updated the uncertainty associated with this measurement by adding those associated with the fitted point-source model and that in the residual map (from which we extracted the flux for RXJ1131) in quadrature (i.e. S_2mm = 0.39+/-0.12 mJy).

The emission in the residual is detected at S/N > 4, which is the significance of the emission in the residual map (Data - Model). This together with the fact that this emission is co-spatial with the lensed arc in the optical image indicates that we detect 2mm continuum towards the background source.

To taken into account the referee's comment, we have updated the 2mm flux in Table 1, and the flux uncertainty to include those from the uv-fitting process.

The term "deblending" is a broadly used in photometry, and it is clear that it refers to the flux subtraction within the same paragraph. We therefore make no changes to the wording.

6) DROP THE SPECTRAL INDEX FOR THE BACKGROUND GALAXY :

Sect.3.3, para.3:

Given the large uncertainty in the residual flux, I don't think
you can derive a 6cm-to-2mm spectral index for the background galaxy.
I suggest dropping that part of the sentence:
(" alpha = -0.345 for the background galaxy ")

If the 2mm continuum flux is synchrotron emission, like the 6cm VLA continuum flux,
then it is probably unrelated to the stellar mass of the quasar host galaxy.

So I don't understand why you use it in sect.4.5, para.4 to fit 
the Bruzual & Charlot stellar templates.

Any comment on the very flat spectrum for the lensing elliptical galaxy ?
Is the 6cm emission mainly synchrotron? 
Is the 2mm emission a mix of synchrotron flux and dust flux ?

# RESPONSE: (explain better)
While the radio-mm spectral slope of -0.35+/-0.21 derived for the background source has a large error bar associated with it due to the large uncertainty on the 2mm flux (0.39+/-0.12 mJy), the fact that it is flatter than the typical slope for pure synchrotron emission indicates that it is a constraint on the fraction of synchrotron emission at 2mm, and that a fraction of the 2mm flux arises from thermal dust and free-free emission.

Flat spectrum of foreground galaxy:
We have also added a statement in the paper, stating that if the background source has a contribution in the point-source model assumed in our deblending procedure, such that its flux density would be higher than derived here (0.39 +/- 0.12 mJy), then the derived spectral index (-0.35) would be even shallower.
In this case, the 2mm flux inferred for the foreground galaxy would lower, and thus steepening its spectral slope. Hence, a contribution from the background source to the point-source model 2mm flux may explain the atypical flat spectrum found for the foreground galaxy.

Decompose synchrotron at 2mm for the background source:
The flatter than pure synchrotron spectral slope found for the background source indicates that the 2mm continuum emission is not dominated by synchrotron emission. Assuming a synchrotron index of alpha = -0.7 and using the 5GHz flux, we expect S_2mmm = 0.122+/-0.004 mJy to be synchrotron emission.
We thus find a flux excess of S_2mm = 0.39 - 0.12 = 0.27 mJy that is coming from thermal dust and free-free emission. We have updated our dust SED fitting using S_2mm = 0.27+/-0.08 mJy and the resulting parameters throughout the paper.

We note that 0.27+/-0.09 mJy is consistent with our previous flux measurement used for SED fitting (0.39+/-0.23 mJy) within the uncertainties, where the uncertainties here include those from absolute flux calibration, therefore changes to the physical parameters derived from SED fitting are only at most a few percent. We have updated the relevant quantities but this does not alter the conclusions of this paper.

7) BETTER LABELS FOR FIGURE 6 AND TABLE 3 :

Figure 6 and Table 3 are hard to understand, because
there are no velocity labels on Fig.6, and because the order 
of the channels in the Figure (most redshifted first, most blueshifted last),
is the opposite of the order in Table 3 
(most blueshifted first, most redshifted last).

The easiest way to fix this would be to :

a) Put some velocity labels on Fig. 6, at least the center velocity
of each of the seven channels.

b) Reverse the order of the channels in Table 3
(change to most redshifted first, most blueshifted last), to 
agree with the order in Fig.6.

# RESPONSE
We have adopted all the suggestions provided by the referee.

8) NEED A TABLE TO SUMMARIZE THE CO RESULTS

There are tables where one can find :

--- the continuum photometry data (Table 1),

--- the model lens parameters (Table 2; please add the mass of the lens to this Table,
even though it is not an independent parameter; otherwise people have to search
for it in the text),

--- the model magnification factors (Table 3; possibly add the derived source offset positions),

--- the fits to the IR continuum SED (Table 4).

But there is no table that gives a quick summary of the 
CO results, --- even though they are a large part of the paper.

I suggest that you add a Table, that as a minimum, might give:

--- the CO(2-1) velocities, linewidths, peak fluxes, line-integrated fluxes,
and spatially integrated fluxes.

--- the CO(3-2) similar parameters (when possible).

--- the L'co luminosities corrected for lensing (RXJ1131 + companion).

--- the M(H2) estimated from the CO luminosity (main source + companion).

The same is true of the derived quantities. 
Possible add another Table summarizing the other derived quantites:

--- R_CO (max) ---From whatever is the best method to estimate it.

--- the M_stellar and SFR ---From the SED fitting.

--- the M_dyn If R = 6 kpc and V = 245 km/s (??)

--- the M_gas / M_dyn ratio

--- the M_gas / M_stars ratio

--- the black hole mass

--- the M_BH / M_bulge ratio.

Otherwise, the reader has to search for all these CO values and other quantities in 
many different parts of the text.

# RESPONSE:
We adopted the good suggestions and have added additional tables summarizing the observed CO line properties and the derived physical parameters.

9. TABLE 1, ERROR BARS :

In Sect.2.1, you say the calibration accuracy at 2mm is 15 per cent;
in Sect.2.3, you say the calibration accuracy at 6cm is 10 per cent.

So in Table 1, the 2mm continuum flux error should include the 
10% measuring error plus the 15% calibration error, which add
as sqrt(sum of squares of percent errors) = 18%.

So S(2mm) = 0.80 +/- 0.14 .

Similarly, the 6cm continuum flux error should be the 3% measuring 
error plus the 10% calibration error (root sum squares) = 10.4%.

So S(6cm) = 0.87 +/- 0.09 .

So you need not quote these fluxes to three decimal places.

Because of these flux errors, you also cannot quote the spectral
index of the foreground galaxy to three decimal places
(Sect. 3.3, last sentence).

# RESPONSE:
It is of standard practice to report flux measurements from interferometric observations excluding those from the absolute flux calibration and state those in the observations section (i.e. Section 2). Throughout the paper, uncertainties from absolute flux calibration are taken into account when deriving the error bars associated with physical parameters. To address the referee's concern, we have added a remark in table 1, and Sec 4.5 para 1 to avoid possible confusion for readers.

We have updated the number of decimals in the spectral indices reported in Sec 3.3, as suggested by the referee.

10. FIGURE 4, CAPTION (VLA MAP) :

Because the lensing galaxy is so strong on the VLA map (Fig.4), one
cannot count the contours on the central source. 
So in the caption, 
it might be good to say how many contours there are up to the peak
(= 22 contours ?), and that this corresponds to a flux of 0.9 mJy/b.

An alternative would be to change the contour step after 3 contours
in the VLA map, so that the higher contours would become visible.

But it is probably easiest to just explain the peak value in 
the figure caption.

# RESPONSE
We have replotted the figure with different contour intervals with thinner lines and added a sentence in the caption to improve the clarity of Fig 4.

11. THE BLR LINES ARE REDSHIFTED, NOT BLUESHIFTED.

Sect.5.3, para.1:
Note that Sluse et al. (2007, their sect.7.2) say the Mg absorption doublet is at 0.654.
Sluse et al (their sect.9, Conclusions) also say the narrow emission lines (the NLR)
are measured to be at 0.654.

So in Sect.5.3,par.1 you should quote them correctly as saying that the first set of lines
is at z = 0.654 (not 0.653).

The 1st set of lines (NLR) is at z = 0.654 (not 0.653).
The 2nd set of lines (BLR) is at z = 0.658 (they are more redshifted).

So this sentence is wrong:

"On the other hand, the second set of lines is blueshifted by 780 km/s,.."

This should read "redshifted", not "blueshifted."

(The 780 km/s is also wrong, 
it should be 700 km/s, see below)

"Using the CO line center redshift as the systemic redshift,
we find that the redshift of the first set is fully consistent
with the "systematic" redshift.

--- Change "systematic" to "systemic".

Sect.5.3, para.3, 2nd sentence from end:

Also wrong, here.

Should be "... our CO observations confirm the redshifted BLR lines."

not "blueshifted".

# RESPONSE:
We have corrected "blueshifted" to "redshifted", updated the velocity offset from 780 km/s to 715 km/s using z = 0.65406, as suggested by the referee.

12. DROP THE "BLACK HOLE RECOIL" SECTION.

You might consider dropping this section entirely.

---The two redshift systems (NLR and BLR) in this quasar have
been known for a long time. 
So the possibility that the NLR-BLR difference is due to a black hole recoil is not 
a new result of your CO-line data, or of your lens modeling, 
which is the main topic of this paper.

---There are many quasars where the BLR redshifts are measured 
to be slightly different from the NLR redshifts.
Sometimes they are more redshifted, sometimes they are more blueshifted.
It is implausible that they all indicate black hole recoils.

---There is a large literature on the redshift differences between
the NLR and BLR lines 
(see e.g., Gaskell, 2009, New Astronomy Reviews, 53, 140; 
Gaskell & Goosmann 2015; Gaskell, Klimek & Nazarova 2007). 
There has been a lot of numerical modeling to explain these differences.

Many simulations favor an inflow of the BLR gas
into the black hole, at a true space velocity of 1000 km/s.

In the BLR, we are actually observing the matter that is flowing into
the black hole. The dominant component in the BLR linewidths is the
rotation of the accretion disk, but there is also an inflow component,
that causes the redshift difference from the NLR lines.

In the simulations, whether the difference is is a blueshift or a redshift
depends on the orientation to the line of sight and the "moving mirror" effect of
the scattering opacity, as well as the clumpiness, etc.

A redshift could result from the disk being more edge-on to our line-of-sight, so 
that clumps in the disk itself block the far side. In this case, we mainly see
the near side, so we see the redshifted inflow component into the black hole.

In the outer parts of the accretion disk, where the density is lower,
there is also a bi-conical outflow, as observed in BAL quasars.

Your phrase 
"...indicating that the BLR of the AGN is "dynamically" offset..."

is not really true. In many current models of the BLR, the region is 
not dynamically offset at all.

The accretion disk rotates around the black hole, and has exactly the same 
cosmological redshift as the host galaxy.

Besides the dominant rotational component, there is also an inflow component, 
that shifts the peak of the line. 
The BLR lines are optically thin, so they go as density squared, 
so the line peak comes from the inner part of the accretion disk, 
where density is highest, 
and where gas flows into the black hole.

For RXJ1131, the picture is further complicated by 
the gravitational lensing, and by the microlensing:

1) MICROLENSING :

Because the NLR is big (100 pc), it is not affected by microlensing.
Because the BLR is small, it is affected by
microlensing in this quasar (Sluse et al.).

It could also be that the microlensing more strongly magnifies
the redshifted part of the BLR lines, 
biasing the apparent redshift to the red side.

2) BLUE SIDE OF THE LINE ATTENUATED BY ABSORPTION :

Another possiblity is that the BLR shows mainly rotation 
in the accretion disk, and also a biconical outflow region
from the outer parts of the accretion disk, as in the BAL phenomenon.

Because of partial, very broad, absorption of the blueshifted flow against the 
AGN continuum, we see stronger emission from the redshifted flow.

Because the system 2 BLR lines in RXJ1131 are very wide (10000 to 20000 km/s), partial
attenuation of the blue sides of the lines could easily shift the apparent 
BLR line centroids toward the red sides of their profiles.

3) UNLIKELY RECOIL VELOCITY :

If the line-of-sight recoil velocity really were 700 km/s, then 
the true space velocity must be even higher, and certainly greater than 
the escape velocity from the host galaxy.

If it is about 1000 km/s, then the black hole moves 1 pc in 1000 years.

So in one rotation time of the host galaxy, 
or in one orbital time of the companion galaxy, 
or in the duration of the starburst, 
or in the duration of the "wet merger" 
(all of these are ~1E8 years), 
the black hole would have moved quite far away from the center of the host galaxy
(10 to 100 kpc !).

This also makes the recoil story implausible, especially for 
many other catalogued quasars and Seyfert galaxies that have differences
between the peaks of the NLR lines and the BLR lines.

# RESPONSE:
We have removed the statement regarding "AGN dynamically offset", as suggested by the referee.

1) & 2)
Our CO measurement provides the most precise value for the velocity offset 
between the BLR and the host galaxy. We thus feel that this is an invaluable addition to the literature. Taking into account the referee's comments, we have updated this section by listing other possible interpretations for the velocity offset (e.g. gas inflows, geometric effects) as part of the discussion, as outlined in detail by the referee.

We agree with the referee that it is unlikely that all offsets in AGN lines are caused by black hole recoil. Considering the referee's comments, we have updated this section to as well discuss other possible scenarios. We only discuss the possibility of a recoil black hole as one of the interpretations for the velocity offset, given that 
    i) the magnitude of the offset in RXJ1131 (>700 km/s) appears to be an outlier in the distribution of offset found for large samples of QSOs (Richards+02, Bonning+07)
    ii) RXJ1131 appears to have a high BH spin parameter (a > 0.9; Reis+14)    iii) recoiling BHs are expected theoretically from hierarchical model of galaxy evolution.

3)
The BH displacement estimated by the referee (~few 10kpc) has been shown to be theoretically possible in the absence of dynamical friction (Bonning+07, Loeb 2007). However, Loeb 2007 has showed that at low ejection speeds of a few hundred km/s, the recoil is expected to be damped by the external gravitational potential of the galaxy, dynamical friction against background stars, dark matter, and gas, thereby limiting the BH displacement and is likely to bring it back to the center of its host galaxy within a few dynamical times (Madau & Quataert 04, Boylan-Kolchin+04, Guedes+11).

We acknowledge that velocity offsets between BLR and NLR lines are quite common in Seyferts and quasars, but the offsets as large as that of RXJ1131 appear to be uncommon according to the distribution of offset among other AGN hosts (Richards+02, Bonning+07). To take into account the referee's comments, we now also discuss other possible scenarios.

TWO FURTHER SUGGESTIONS FOR SHORTENING THE PAPER:

b) ---Drop the detailed description of the best-fit procedures
(The Markov chain Monte Carlo paragraph in Sect.4.1, paragraph 3).

c) ---Drop the details of the SED fitting procedure in Sect.4.5, 
(especially in the 2nd half of paragraph 1: 
"We impose a uniform prior"...and all the text that follows.)

Both of these sections are very well-written, but will probably be
completely incomprehensible to many readers (including me).

I suggest that you just say that you did best fits to model the lens distribution
and best fits to the SED. 
Simply give a typical reference to the Markov chain, Monte Carlo method.

# RESPONSE:
The details of the model (e.g. how we reconstruct the likelihood functions) are described in other papers, which we reference to in the paper for brevity.

However, the description of our MCMC sampling has been retained since stating the priors and the methods to ensure convergence are necessary to make results reproducible, and provide crucial information for readers to decide how trust-worthy our inference are from a statistical viewpoint. Explicitly stating the priors is important to demonstrate that our choice are reasonable and physically motivated since they could affect the marginal posteriors strongly, and thus the inferred physical parameters. We therefore decide to keep them in the lens modeling (Sec 4.1) and SED fitting sections (Sec 4.5).

13. FIX THE CONCLUSIONS

Sect.6, para.2: NUMBERS SLIGHTLY INCONSISTENT:

You derive M_gas = 1.38E10 Msun, 
and also M_dyn = 8E10 Msun.

So the ratio f_gas = M_gas / M_dyn = 16.6 percent, not 19 percent.

(This also has to be corrected in the Abstract, and in Sect.5.1.2, para.1).

Sect.6, para.4: DROP THE BLACK HOLE RECOIL STORY.

"...but the optical BLR lines are blueshifted by 780 km/s."

Also wrong here. The BLR lines are redshifted, not blueshifted :

BLR = 0.658 (Sect.5.3, para.1);

CO = 0.6541, NLR lines = 0.654 (Sect.5.3).

If CO(2-1) were at z = 0.658, its frequency would be 139.046 GHz.
True CO(2-1) is at z = 0.6541, so its frequency is 139.373 GHz.

So the apparent z = 0.658 is 327 MHz lower in frequency than systemic,
which corresponds to a Doppler velocity = 327 x 2.151 km/s/MHz,

or V_doppler = 703 km/s higher in velocity (more redshifted)
than the true CO(2-1) velocity centroid at z = 0.6541.

So the BLR lines should be "redshifted by 700 km/s" (not 780 km/s).

OMIT THE REST OF THIS PARAGRAPH,

starting with "This implies that the AGN is offset from the dynamical
center of its host galaxy." all the way to the end to the paragraph.

This is not true. The AGN could be at the center of the galaxy,
and the Doppler shift could just be a radial component (inflow),
superposed on the very fast rotational velocity of the accretion disk.

# RESPONSE:
We updated f_gas and f_mol using M_gas and M_dyn of both RXJ1131+companion , i.e., as a merger system.

For the other comments, please refer to our response to point 11 and 12, where we describe how we took them into account.

14. CHANGE FIG.9, (Left Panel) : DROP THE VELOCITY MOMENT MAP (SAME AS FIG.2)

I found Fig.9 (left) very confusing.

You should drop the velocity moment map (its the same map as in Fig.2, anyway), 
and instead, only plot the little crosses
in the source plane, on a more blown-up RA-Dec scale, so the crosses would be more visible.

The diagram is confusing because the moment map is the final observed result 
in the sky plane, after lensing, 
with lensed positions quite different from the original positions in the source plane.

Furthermore, the moment map has a weighting of velocities 
that is biased by the differential lensing, so that the 
red velocities get higher weighting than the blue velocities.

All of this distorted image in the sky plane gets even further distorted 
by convolution with the quite large 4.4" x 2.0" synthesized beam.

The little crosses, however, are model positions (not observed data, unlike the moment map)
in the source plane (not the sky plane),
before lensing, without any intensity weighting, 
and before any convolution with the instrumental beam.

Because this figure is about the result of the best-fit lens models, and because the 
little crosses are the basis for the model p-v curve in the right panel (and not the 
contours on the moment map), 
it would be better to concentrate on the model results in the source plane,
and drop the confusing moment map in the image plane, and drop the synthesized beam ellipse.

That way, on a blown-up diagram, you would have room to put velocity labels on each of the
little crosses, so they could be more easily related to the p-v diagram in the right panel.

Could another interpretation be that instead of tracing a monotonic p-v diagram,
there are, within the errors, just two little clusters of points in the left panel?
---The upper (red wing) cluster and the lower (blue wing) cluster?
Within the errors, the 3 crosses in the red clump almost agree in position,
and the 4 crosses in the blue clump almost agree in position, as well.

The separation between the centroids of the two clusters of points is 1 arcsec, or 7 kpc.
So if this is a central disk, then within the errors, its radius is only 3.5 kpc, not 
the 6 kpc that you quote in many places in the text.

Why do these two clusters of points in Fig.9 (left) not appear in your
p-v diagram in Fig.9 (right)?
In the right panel, why do all the little crosses seem to be evenly spaced?
Is the right panel based on an additional model of a rotating disk, rather than 
the lens model points in Fig.9 (left)?

# RESPONSE:
Indeed, the plot on the left is to demonstrate that while the observed 1st moment map is distorted due to a combination of differential lensing and beam smearing effect, our dynamical lens modeling approach, which finds the best-fit models by taking into account the beam convolution indirectly, has enabled us to reconstruct the source-plane kinematic components. Therefore, we show the source-plane positions atop of the observed 1st moment map. To address the referee's concern, we have added some additional description in the caption to clarify what is being shown in the left panel.

The reconstructed source-plane positions display a clear velocity gradient. Taken together with previous results from literature (C06) and the intrinsically symmetric line profile (Fig. 5), we carried out dynamical modeling in Sec 4.3 assuming RXJ1131 is a disk galaxy.

PV: 
We have detailed the reasons for why we interpret RXJ1131 as a disk galaxy in Sec 4.1.1. Under this assumption, the markers on the right plot are based on extracting along a fitted major axis of a disk (see first paragraph of Sec 4.3).

To address the referee's concern and to improve the clarity of this section, we have re-iterate in the beginning of Sec 4.3 that we assume RXJ1131 to be a disk galaxy for dynamical modeling.

Size:
As explained in para. 3 of Sec 4.3, we use the maximum observed velocity (i.e.  corresponding to the redmost channel along the disk major axis) at 6+/-3kpc as a proxy for the rotation velocity. Thus the quoted CO disk radius corresponds to the extent of this velocity component, as measured from the centroid position at zero velocity.

Clumps:
Given that RXJ1131 shows a symmetric double-horned line profile (Fig. 5) and a disk-like morphology with a Sersic index of n=1 in the source-plane of optical-NIR emission (C06), the source-plane velocity gradient found by lens modeling the CO emission (i.e., based on the source-plane positions of different velocity components as shown in the left panel of Fig 9) suggest that RXJ1131 is more likely to be a single rotating disk rather than two kinematically-ordered clumps.